### 2013-03-28 doc/sosp13/evaluation.tex
finished eval structure

\subsection{Basic Grappa mechanisms}

\subsection{Fast User-Level Context Switching }

% More specifically, each context is cache lines: 1 for the worker struct and 3
% for stack data. So the are so 8 cacheline transfers per context switch (write
% the old context, read the next in). The bandwidth of single socket is 270M cache lines/s.
% 
% This asymptotically approaches
% 33.75Mcontexts/s as you increase the number of contexts (fewer and fewer in
% L3). Note that only 4+ cores reaches full rate because these westmere chips
% are balanced to not achieve full off-chip bandwidth until 4 cores.

@@

@@

>>>>

\subsection{Basic \Grappa Mechanisms}

\paragraph{User-level context switching.}
As discussed earlier, fast context switching is at the heart of \Grappa's
latency tolerance abilities. We assess context switch overheads using a simple
microbenchmark that has a configurable number of workers, where each worker 
just increments values in a large array.

\begin{figure}[ht]
    \begin{center}
      \includegraphics[width=0.5\textwidth]{figs/context_switch_time.pdf}
%      \includegraphics[width=0.5\textwidth]{figs/context_switch_bw.pdf}
    \end{center}
    \caption{Average context switch time with 1 active, with and without prefetching.}
    \label{fig:context-switch-exp}
\end{figure}

Figure~\ref{fig:context-switch-exp} shows the average context switch time as
the number of workers grow. There are several important takeaway points form
these data. Context switch time for the number of workers we use currently
($~$1K) is on the order of 50 ns; and prefetching has significant impact on
context switch time. And ontext switch time does not grow significantly with
the number of workers --- even at close to one half million workers context
switch is around 75ns.

We also measured (not in the plot) the \emph{rate} of context switch for all
cores in a node, which showed that our rate is limited by off-chip memory
bandwidth. Each context is 4 cache lines (1 for the worker struct and 3 for
stack data), leading to 8 cache line transfers per context switch
(write the previous context, read the next in). The off-chip bandwidth of a
single socket in our system is 270M cache lines per second, which implies that
we can sustain at most 34M context switches per second per socket, which is
almost exactly what we sustain.

In summary, our context switch engine is able to efficiently sustain very high concurrency and we will show later, the amount of concurrency sustained is sufficient the latencies \Grappa needs to hide.

\paragraph{Global memory and communication.} We measure the performance of
\Grappa's global memory and communication layers using a faithful
implementation of the giga updates per second (GUPS) benchmark.
Read-modify-write updates are dispatched at random to a large array. This
benchmark stresses the communication layer of \Grappa separately from the
scheduler, because only a single worker is used per system node. Figure~\ref{fig:grappa-gups} shows that Grappa is able to sustain on the order of billions of updates per second, which is enough to \TODO{...}.

\begin{figure}[ht]
    \begin{center}
%      \includegraphics[width=0.5\textwidth]{figs/grappa-gups.pdf}
    \end{center}
    \caption{GUPS (giga updates per second) for \Grappa as the number of nodes grows.}
    \label{fig:grappa-gups}
\end{figure}

\paragraph{Putting it all together with Unbalanced Tree Search (UTS).}
We now show the overall performance of \Grappa running UTS. Figure~\ref{fig:grappa-uts}. The point of this experiment is to show that \Grappa's context switching and communication layers can in fact be used together to run an irregular application efficiently. Looking at the performance as the number of nodes grows shows \TODO{...}.

\begin{figure}[ht]
    \begin{center}
%      \includegraphics[width=0.5\textwidth]{figs/grappa-uts.pdf}
    \end{center}
    \caption{Vertices per second in UTS on \Grappa as the number of nodes grows.}
    \label{fig:grappa-uts}
\end{figure}

\subsection{Comparing \Grappa to Other Systems}

In order to ground \Grappa's performance, we compare it with XMT running BFS,
PageRank, IntSort, GUPS and UTS. Since XMT is a different hardware platform,
we also compare \Grappa with hand-tuned MPI versions of BFS and GUPS running
on the same hardware. Finally, we also compare it with UTS written for UPC. We
run all experiments with 64 nodes.

\begin{figure}[ht]
    \begin{center}
%      \includegraphics[width=0.5\textwidth]{figs/grappa-uts.pdf}
    \end{center}
    \caption{Comparing \Grappa with XMT, hand-tuned MPI and UPC.}
%bar chart, one set of bars per benchmark, one bar per system. runtime normalized to grappa.
    \label{fig:grappa-comparisons}
\end{figure}

Figure~\ref{fig:grappa-comparisons} shows the results. Overall, \Grappa is
within striking distance of XMT and UPC performance. However, it underperforms
compared to hand-tuned MPI. Nevertheless, this comparison needs to be taken
with a grain of salt for several reasons. First, we have not spent as much
time tuning \Grappa's implementations. This is supported by inspecting the BFS
MPI implementation we used, which clearly shows that it employs a lot of
algorithmic-specific optimizations. In fact, some of these optimizations
resemble some of what \Grappa does automatically, like message aggregation.
Second, the GUPS results presented earlier suggests that \Grappa performance
can do a lot better. \TODO{update this once we have more results. }

\subsection{Characterization}

\paragraph{Where execution time goes.}

\paragraph{Message size and latency.}
Distribution if possible. With and without aggregation.

\paragraph{Frequency of remote requests.}

\paragraph{Work stealing behavior.}
Performance loss of not having stealing. Ratio of \# of steals over total \# of task spawns. UTS needs stealing to work.

@@

\comment{

@@

}

### 2013-03-28 doc/sosp13/method.tex
finished eval structure

\noindent{\bf Context-switch:} This simple microbenchmark is used to explore context switch overhead.  A configurable number of workers are used to increment the values in a large array.  This benchmark stresses the scheduler inside of Grappa.

\noindent{\bf GUPS:} This is a faithful implementation of the
giga-updates-per-second benchmark. Read-modify-write updates are dispatched at
random to a large array. This benchmark stresses the networking layer of
Grappa separately from the scheduler, because only a single worker is used per
system node.

\noindent{\bf Unbalanced tree search in-memory (UTS-Mem):} Unbalanced Tree Search
(UTS) is a benchmark for evaluating the programmability and performance of
systems for parallel applications that require dynamic load
balancing~\cite{UTS}. It involves traversing an unbalanced implicit tree: at
each vertex, its number of children is sampled from some probability
distribution, and this number of new nodes are added to a work queue to be
visited. While this benchmark captures irregular, dynamic \emph{computation,}
we actually want to evaluate performance of algorithms with irregular
\emph{memory\/} access patterns. Thus we augment UTS by using the existing
traversal code to create a large tree in memory, and then we traverse the
in-memory tree. We call this benchmark UTS-Mem, and the timed portion is this
traversal of the in-memory tree. This in-memory traversal has no knowledge of
the tree structure beforehand.

\noindent{\bf Breadth-first-search (BFS):} This is the primary kernel for the
Graph500 benchmark and is what currently determines the ranking of machines on
the Graph500 list~\cite{graph500list}. As a whole, the Graph500 benchmark
suite is designed to bring the focus of system design on data-intensive
workloads, particularly large-scale graph analysis problems, that are
important among cybersecurity, informatics, and network understanding
workloads. The BFS benchmark builds a search tree containing parent nodes for
each traversed vertex during the search. While this is a relatively simple
problem to solve, it exercises the random-access and fine-grained
synchronization capabilities of a system as well as being a primitive in many
other graph algorithms. Performance is measured in \emph{traversed edges per
second\/} (TEPS), where the number of edges is the edges making up the
generated BFS tree. One of the reference implementations of Graph500 BFS is
for the XMT; this code fails to scale past 16 XMT processors because it does
not expose enough parallelism, so we modified the code to use a recursive loop
decomposition similar to Grappa's. We compare this modified version against a
straightforward Grappa implementation. We do not employ algorithmic
improvements, though there are
many~\cite{Beamer:Graph500,Yoo:FixedPointGraph500}.

\noindent{\bf IntSort:} \TODO{Write me Brandon Holt!}

\noindent{\bf PageRank:} \TODO{Write me Brandon Myers!}

>>>>

%%% Moved the below to the evaluation sec
% \vspace{0.5ex}
% \noindent{\bf Context-switch:} This simple microbenchmark is used to explore context switch overhead.  A configurable number of workers are used to increment the values in a large array.  This benchmark stresses the scheduler inside of Grappa.
% 
% \vspace{0.5ex}
% \noindent{\bf GUPS:} This is a faithful implementation of the
% giga-updates-per-second benchmark. Read-modify-write updates are dispatched at
% random to a large array. This benchmark stresses the networking layer of
% Grappa separately from the scheduler, because only a single worker is used per
% system node.

\vspace{0.5ex}
\noindent{\bf Unbalanced tree search in-memory (UTS-Mem):} Unbalanced Tree Search
(UTS) is a benchmark for evaluating the programmability and performance of
systems for parallel applications that require dynamic load
balancing~\cite{UTS}. It involves traversing an unbalanced implicit tree: at
each vertex, its number of children is sampled from some probability
distribution, and this number of new nodes are added to a work queue to be
visited. While this benchmark captures irregular, dynamic \emph{computation,}
we actually want to evaluate performance of algorithms with irregular
\emph{memory\/} access patterns. Thus we augment UTS by using the existing
traversal code to create a large tree in memory, and then we traverse the
in-memory tree. We call this benchmark UTS-Mem, and the timed portion is this
traversal of the in-memory tree. This in-memory traversal has no knowledge of
the tree structure beforehand.

\vspace{0.5ex}
\noindent{\bf Breadth-first-search (BFS):} This is the primary kernel for the
Graph500 benchmark and is what currently determines the ranking of machines on
the Graph500 list~\cite{graph500list}. As a whole, the Graph500 benchmark
suite is designed to bring the focus of system design on data-intensive
workloads, particularly large-scale graph analysis problems, that are
important among cybersecurity, informatics, and network understanding
workloads. The BFS benchmark builds a search tree containing parent nodes for
each traversed vertex during the search. While this is a relatively simple
problem to solve, it exercises the random-access and fine-grained
synchronization capabilities of a system as well as being a primitive in many
other graph algorithms. Performance is measured in \emph{traversed edges per
second\/} (TEPS), where the number of edges is the edges making up the
generated BFS tree. One of the reference implementations of Graph500 BFS is
for the XMT; this code fails to scale past 16 XMT processors because it does
not expose enough parallelism, so we modified the code to use a recursive loop
decomposition similar to Grappa's. We compare this modified version against a
straightforward Grappa implementation. We do not employ algorithmic
improvements, though there are
many~\cite{Beamer:Graph500,Yoo:FixedPointGraph500}.

\vspace{0.5ex}
\noindent{\bf IntSort:} \TODO{Write me Brandon Holt!}

\vspace{0.5ex}
\noindent{\bf PageRank:} \TODO{Write me Brandon Myers!}
